---
title: "Trajectory analysis of Turkey vultures"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Trajectory analysis of Turkey vultures}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, fig.width = 7, fig.height = 7,
  comment = "#>"
)
if (Sys.info()["user"] != "bart") {
  if (Sys.getenv("MBPWD") != "") {
    options(keyring_backend = "env")
    move2::movebank_store_credentials("move2_user", Sys.getenv("MBPWD"))
  } else {
    knitr::opts_chunk$set(eval = FALSE)
  }
}
Sys.setlocale("LC_TIME", "en_US.utf8")
```

## Retrieve data

```{r setup}
library(move2)
vulture_data <-
  movebank_download_study("Turkey vultures in North and South America")
vulture_data
```

In this case some tracks have very long time gaps, to prevent distant points to
be connected by a line we split those tracks by creating a new id.

```{r}
library(dplyr, quietly = TRUE)
library(ggplot2, quietly = TRUE)
library(rnaturalearth, quietly = TRUE)
library(units, quietly = TRUE)
vulture_lines <- vulture_data %>%
  mutate_track_data(name = individual_local_identifier) %>%
  mutate(
    large_gaps = !(mt_time_lags(.) < set_units(1500, "h") |
      is.na(mt_time_lags(.))),
    track_sub_id = cumsum(lag(large_gaps, default = FALSE)),
    new_track_id = paste(mt_track_id(.), track_sub_id)
  ) %>%
  mt_set_track_id("new_track_id") %>%
  mt_track_lines()
ggplot() +
  geom_sf(data = ne_coastline(returnclass = "sf", 50)) +
  theme_linedraw() +
  geom_sf(
    data = vulture_lines,
    aes(color = name)
  ) +
  coord_sf(
    crs = sf::st_crs("+proj=aeqd +lon_0=-83 +lat_0=8 +units=km"),
    xlim = c(-3500, 3800), ylim = c(-4980, 4900)
  )
```

## Categorize seasons

```{r}
library(magrittr, quietly = TRUE)
library(lubridate, quietly = TRUE)
vulture_data <- vulture_data %>% mutate(
  month = month(mt_time(), label = TRUE, abbr = FALSE),
  season = recode_factor(month,
    January = "Wintering", February = "Wintering",
    March = "Wintering", April = "North migration",
    May = "North migration", June = "Breeding",
    July = "Breeding", August = "Breeding",
    September = "South migration", October = "South migration",
    November = "Wintering", December = "Wintering"
  ),
  # Here we change season to NA if the next location is either from
  # a different track or season
  season = if_else(season == lead(season, 1) &
    mt_track_id() == lead(mt_track_id(), 1),
  season, NA
  )
)
```

Annotate speed and azimuth to the trajectory.

```{r}
vulture_data <- vulture_data %>% mutate(azimuth = mt_azimuth(.), speed = mt_speed(.))
```
## Seasonal distribution per individual
```{r fig.height=6}
library(circular, quietly = TRUE)
vulture_azimuth_distributions <- vulture_data %>%
  filter(speed > set_units(2, "m/s"), !is.na(season)) %>%
  group_by(season, track_id = mt_track_id()) %>%
  filter(n() > 50) %>%
  summarise(azimuth_distribution = list(density(
    as.circular(
      drop_units(set_units(
        azimuth,
        "degrees"
      )),
      units = "degrees",
      modulo = "asis",
      zero = 0,
      template = "geographic", rotation = "clock", type = "angles"
    ),
    bw = 180, kernel = "vonmises"
  )))

# Load purrr for map function
library(purrr, quietly = TRUE)
# Load tidy r for unnest function
library(tidyr, quietly = TRUE)
vulture_azimuth_distributions %>%
  mutate(
    x = map(azimuth_distribution, ~ .$x),
    y = map(azimuth_distribution, ~ .$y)
  ) %>%
  select(-azimuth_distribution) %>%
  unnest(c(x, y)) %>%
  ggplot() +
  geom_path(aes(x = x, y = y, color = season)) +
  coord_polar(start = pi / 2) +
  theme_linedraw() +
  facet_wrap(~ factor(track_id)) +
  scale_x_continuous(
    name = NULL, breaks = (-2:1) * 90,
    labels = c("S", "W", "N", "E")
  ) +
  scale_y_continuous(name = NULL, limits = c(-0.8, 1.0), expand = c(0L, 0L))
```

## Individual trajectory

```{r, fig.height=5}
leo <- vulture_data |>
  filter_track_data(individual_local_identifier == "Leo") |>
  mutate(speed_categorical = cut(speed, breaks = c(2, 5, 10, 15, 35)))
leo |> ggplot(aes(x = azimuth, y = speed)) +
  geom_point() +
  scale_x_units(unit = "degrees", breaks = -2:2 * 90, expand = c(0L, 0L)) +
  theme_linedraw()
```
```{r}
leo |>
  filter(speed > set_units(2L, "m/s"), !is.na(season)) |>
  ggplot() +
  coord_polar(start = pi) +
  geom_histogram(
    aes(
      x = set_units(azimuth, "degrees"),
      fill = speed_categorical
    ),
    breaks = set_units(seq(-180L, 180L, by = 10L), "degrees"),
    position = position_stack(reverse = TRUE)
  ) +
  scale_x_units(
    name = NULL,
    limits = set_units(c(-180L, 180L), "degrees"),
    breaks = (-2L:2L) * 90L
  ) +
  facet_wrap(~season) +
  scale_fill_ordinal("Speed") +
  theme_linedraw()
```

Plot turn angle distribution.

```{r, fig.height=4, fig.width=5}
pi_r <- set_units(pi, "rad")
leo %>%
  mutate(turnangle = mt_turnangle(.)) %>%
  filter(speed > set_units(2L, "m/s"), lag(speed, 1L) > set_units(2L, "m/s")) %>%
  ggplot() +
  geom_histogram(
    aes(
      x = turnangle,
      fill = speed_categorical
    ),
    position = position_stack(reverse = TRUE)
  ) +
  scale_fill_ordinal("Speed") +
  coord_polar(start = pi) +
  scale_x_units(limits = c(-pi_r, pi_r), name = NULL) +
  scale_y_continuous(limits = c(-500L, 650L), breaks = c(0L, 250L, 500L)) +
  theme_linedraw()
```

## Net displacement

Here we visualize the distance to the first location of each trajectory.

```{r, fig.height=5}
vulture_data <- vulture_data %>%
  group_by(mt_track_id()) %>%
  mutate(displacement = c(st_distance(
    !!!syms(attr(., "sf_column")),
    (!!!syms(attr(., "sf_column")))[row_number() == 1]
  )))

vulture_data %>% ggplot() +
  geom_line(aes(
    x = timestamp,
    y = set_units(displacement, "km"),
    color = individual_local_identifier
  )) +
  ylab("Distance from start") +
  theme_linedraw()
```
